---
title: "Cremella et al. 2022 - processing code"
knit: (function(input_file, encoding) {
    out_dir <- 'docs';
    rmarkdown::render(input_file,
      encoding=encoding,
      output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
author: "Bruno Cremella"
date: "06/06/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# AQUATIC PARTICULATE ABSORPTION COEFFICIENT COMBINING EXTRACTION AND BLEACHING 
# OPTIMIZED FOR INLAND WATERS

## Cremella, B., Bélanger, S., and Huot, Y.
Limnology and Oceanography: Methods, June 2022.
<https://doi.org/10.1002/lom3.10497>

This is the processing code for the corresponding L&O:M article.

## 1	RAW DATA PROCESSING

* Raw absorbances are loaded as matrices of n wavelength rows X m filters (filters named as lake ID underscore replicate). Filter properties are loaded too (as a matrix of n properties X m filters).   Properties include lake ID, replicate (A, B, or X for blanks), filtered volume and filtration area diameter. 
* Deionized water blanks were filtered in random lakes, and were treated normally (no treatment, methanol treatment, and bleaching treatment measurements). Blanks are averaged and the resulting blank mean absorbance is subtracted to the corresponding batch of samples (one mean per treatment per year). 
* Mean NIR absorption for null point correction was calculated between 750 and 800 nm (rows 1:50), since a hump potentially of long wavelength chlorophyll complexes that absorb in the 700-720 nm range (sometimes present in cyanobacteria) was found in some samples. 
* Absorption coefficients are calculated according to Stramski et al. 2015 formula, using filtered volume and filter diameter from filter properties. Exponent was changed to 1 to account for negative absorption.

First, load the matrices and filter properties
```{r, eval=FALSE}
#setwd("C:/Users/bruno/OneDrive - USherbrooke/PhD thesis analysis 2021/Extra.aPhy method subchapter/For github")
#----Step 1 ---- Raw data processing-----
#------Step 1.1: Raw data loading and preparation----
wholepabs=read.csv2("whole.csv") #untreated PABS absorbances matrix 
extracted=read.csv2("extracted.csv") #methanol treated absorbance matrix
bleached=read.csv2("bleached.csv") # bleached treated absorbance matrix
filterprops=read.csv2("properties.csv") #filter properties: filtered volume
```
Deionized water blanks were filtered in random lakes, and were treated normally (no treatment, methanol treatment, and bleaching treatment measurements). Blanks are averaged and the resulting blank mean absorbance is subtracted to the corresponding batch of samples (one mean per treatment per year). 
```{r, eval=FALSE}
#------Step 1.2: Field blank subtraction ----
library(dplyr);library(janitor);library(inspectr)
filterprops.fieldblanks=filterprops[,grepl("X",filterprops[3,])] 
#row 3 contain replicate, select filters
#row 9 contains sampling period (for Lake Pulse: year)
wholepabs.fieldblanks=rbind(as.character(filterprops.fieldblanks[9,]),
  wholepabs[,grepl("X",filterprops[3,])])
#
extracted.fieldblanks=rbind(as.character(filterprops.fieldblanks[9,]),
 extracted[,grepl("X",filterprops[3,])])
#
bleached.fieldblanks=rbind(as.character(filterprops.fieldblanks[9,]),
 bleached[,grepl("X",filterprops[3,])])
#plots indicates that blanks average per year are ok
#proceed to calculate mean blank per year per treatment
mean.wholepabs.yearblank=wholepabs.fieldblanks%>%t%>%as.data.frame%>%
  dplyr::rename(year=V1)%>%group_by(year)%>%mutate_all(funs(as.numeric))%>%
  summarise_all(funs(mean),na.rm=T)%>%t%>%as.data.frame%>%
  row_to_names(row_number = 1)
mean.wholepabs.yearblank=cbind(wholepabs[,1],mean.wholepabs.yearblank)
#
mean.extracted.yearblank=extracted.fieldblanks%>%t%>%as.data.frame%>%
  dplyr::rename(year=V1)%>%group_by(year)%>%mutate_all(funs(as.numeric))%>%
  summarise_all(funs(mean),na.rm=T)%>%t%>%as.data.frame%>%
  row_to_names(row_number = 1)
mean.extracted.yearblank=cbind(wholepabs[,1],mean.extracted.yearblank)
#
mean.bleached.yearblank=bleached.fieldblanks%>%t%>%as.data.frame%>%
  dplyr::rename(year=V1)%>%group_by(year)%>%mutate_all(funs(as.numeric))%>%
  summarise_all(funs(mean),na.rm=T)%>%t%>%as.data.frame%>%
  row_to_names(row_number = 1)
mean.bleached.yearblank=cbind(wholepabs[,1],mean.bleached.yearblank)

#Remove blank filters
filterprops.nox=filterprops[,!grepl("X",filterprops[3,])] #row 3 is replicate
wholepabs.nox=rbind(as.character(filterprops.nox[9,]),
                    wholepabs[,!grepl("X",filterprops[3,])]) #row 9 is year
extracted.nox=rbind(as.character(filterprops.nox[9,]),
                    extracted[,!grepl("X",filterprops[3,])])
bleached.nox=rbind(as.character(filterprops.nox[9,]),
                   bleached[,!grepl("X",filterprops[3,])])
#Blank Subtraction
wholepabs.blankcor=data.frame(mapply(wholepabs.nox[-1,],FUN=as.numeric)-
 mapply((mean.wholepabs.yearblank[,match(wholepabs.nox[1,],
 names(mean.wholepabs.yearblank),nomatch=1)]),FUN=as.numeric))
wholepabs.blankcor[,1]=wholepabs[,1]
extracted.blankcor=data.frame(mapply(extracted.nox[-1,],FUN=as.numeric)-
 mapply((mean.extracted.yearblank[,match(extracted.nox[1,],
 names(mean.extracted.yearblank),nomatch=1)]),FUN=as.numeric))
extracted.blankcor[,1]=extracted[,1]
bleached.blankcor=data.frame(mapply(bleached.nox[-1,],FUN=as.numeric)-
 mapply((mean.bleached.yearblank[,match(bleached.nox[1,],
 names(mean.bleached.yearblank),nomatch=1)]),FUN=as.numeric))
bleached.blankcor[,1]=bleached[,1]

```
Mean NIR absorption for null point correction was calculated between 750 and 800 nm (rows 1:50), since a hump potentially of long wavelength chlorophyll complexes that absorb in the 700-720 nm range (sometimes present in cyanobacteria) was found in some samples.  
```{r, eval=FALSE}
#----- Step 1.3 Pseudo-null point correction-----
#CHECK THAT ALL FILTER HAVE VOLUME IN THEM, IF NOT CHANGE IT TO 999999
#mean NIR was calculated between 750 and 800 (rows 1:50)
filterprops.nox=rbind(filterprops.nox,c("NIR below 0",
                                        rep(NA,n=(length(filterprops.nox)-1))))
wholepabs.fix=wholepabs.blankcor
for (i in (2:ncol((wholepabs.blankcor)))) {
  if (mean(wholepabs.blankcor[1:50,i])<0) {wholepabs.fix[,i]= (wholepabs.blankcor[,i]-mean(wholepabs.blankcor[1:50,i])) 
  filterprops.nox[10,i]="PABSNeg" }
  else {(wholepabs.fix[,i]=wholepabs.blankcor[,i])
    filterprops.nox[10,i]="PABSOK" } }
extracted.fix=extracted.blankcor
for (i in (2:length(names(extracted.blankcor)))) {
  if(mean(extracted.blankcor[1:50,i])<0) {extracted.fix[,i]=(extracted.blankcor[,i]-mean(extracted.blankcor[1:50,i])) 
  filterprops.nox[10,i]=paste(filterprops.nox[10,i], "ExtNeg",sep="_")}
  else {extracted.fix[,i]=(extracted.blankcor[,i])
  filterprops.nox[10,i]=paste(filterprops.nox[10,i], "ExtOK",sep="_")}}
bleached.fix=bleached.blankcor
for (i in (2:length(names(bleached.blankcor)))) {
  if(mean(bleached.blankcor[1:50,i])<0) {bleached.fix[,i]=(bleached.blankcor[,i]-mean(bleached.blankcor[1:50,i])) 
  filterprops.nox[10,i]=paste(filterprops.nox[10,i], "BleNeg",sep="_") }
  else { bleached.fix[,i]=(bleached.blankcor[,i])
  filterprops.nox[10,i]=paste(filterprops.nox[10,i], "BleOK",sep="_")     }}
#filterprops.nox row 10 is the null point correction flag
#unt. filters with  negative NIR 
length(grep("PABSNeg",filterprops.nox[10,-1]))
#post extraction filters with negative NIR
length(grep("ExtNeg",filterprops.nox[10,-1]))
#post bleached filters with negative NIR
length(grep("BleNeg",filterprops.nox[10,-1]))
#----- Step 1.4 NAP-ABS higher than PABS, pseudonull point correction-----
filterprops.nox=rbind(filterprops.nox,c("higher NIR processed",
                                        rep(NA,n=(length(filterprops.nox)-1))))
#Differences between NIR absorption between treatments
extracted.nirfix=extracted.fix
differences.exnir=colMeans(extracted.fix[(1:50),])-
  colMeans(wholepabs.fix[(1:50),])
for (i in (2:ncol(extracted.fix))) {
if (differences.exnir[i]>0) 
{extracted.nirfix[,i]=  extracted.fix[,i]-as.numeric(differences.exnir[i])
filterprops.nox[11,i]="ExtHigher"} # Extraction NIR is higher than untreated NIR
else {extracted.nirfix[,i]=  extracted.fix[,i]
filterprops.nox[11,i]="ExtOK"}   } # NIR is OK
bleached.nirfix=bleached.fix
differences.bnir=colMeans(bleached.fix[(1:50),])-
  colMeans(extracted.fix[(1:50),])
for (i in (2:ncol(bleached.fix))) {
  if (differences.bnir[i]>0) 
  {bleached.nirfix[,i]=  bleached.fix[,i]-as.numeric(differences.bnir[i])
  filterprops.nox[11,i]=paste(filterprops.nox[11,i], "BleHigher",sep="_")  } 
  # Bleached  NIR is higher than extracted  NIR
  else {bleached.nirfix[,i]=  bleached.fix[,i]
  filterprops.nox[11,i]=paste(filterprops.nox[11,i], "BleOK",sep="_")   }   } 
# Bleached  NIR is OK
```
Absorption coefficients are calculated according to Stramski et al. 2015 formula, using filtered volume and filter diameter from filter properties. Exponent was changed to 1 to account for negative absorption.

```{r, eval=FALSE}
#----- Step 1.5 Absorption coefficient transformation ----
#Stramski et al. 2015 pathlength amplification correction
#For negative absorption, change the exponent to 1
wholepabs.acoef=wholepabs.fix;wholepabs.acoef[,-1]<- NA
for (i in (2:ncol(wholepabs.fix))) {
 for (j in (1:nrow(wholepabs.fix))) {
   wholepabs.acoef[j,i]= 
  if(wholepabs.fix[j,i]<0) 
    ((pi*((1e-3*as.numeric(filterprops.nox[4,i])/2)^2))*2.303*0.323*
       wholepabs.fix[j,i])/(1e-6*as.numeric(filterprops.nox[5,i])) 
  else  ((pi*((1e-3*as.numeric(filterprops.nox[4,i])/2)^2))*2.303*0.323*
        (wholepabs.fix[j,i]^1.0867))/(1e-6*as.numeric(filterprops.nox[5,i])) }}

extracted.acoef=extracted.nirfix;extracted.acoef[,-1]<- NA
for (i in (2:ncol(extracted.nirfix))) {
  for (j in (1:nrow(extracted.nirfix))) {
    extracted.acoef[j,i]= 
      if(extracted.nirfix[j,i]<0) 
        ((pi*((1e-3*as.numeric(filterprops.nox[4,i])/2)^2))*2.303*0.323*
           extracted.nirfix[j,i])/(1e-6*as.numeric(filterprops.nox[5,i])) 
    else  ((pi*((1e-3*as.numeric(filterprops.nox[4,i])/2)^2))*2.303*0.323*
    (extracted.nirfix[j,i]^1.0867))/(1e-6*as.numeric(filterprops.nox[5,i])) }}

bleached.acoef=bleached.nirfix;bleached.acoef[,-1]<- NA
for (i in (2:ncol(bleached.nirfix))) {
  for (j in (1:nrow(bleached.nirfix))) {
    bleached.acoef[j,i]= 
      if(bleached.nirfix[j,i]<0) 
        ((pi*((1e-3*as.numeric(filterprops.nox[4,i])/2)^2))*2.303*0.323*
           bleached.nirfix[j,i])/(1e-6*as.numeric(filterprops.nox[5,i])) 
    else  ((pi*((1e-3*as.numeric(filterprops.nox[4,i])/2)^2))*2.303*0.323*
      (bleached.nirfix[j,i]^1.0867))/(1e-6*as.numeric(filterprops.nox[5,i])) }}

#----- Step 1.6 Absorption coefficient export ----
write.csv(wholepabs.acoef,"LP.wholepabs.acoef.csv")
write.csv(extracted.acoef,"LP.extracted.acoef.csv")
write.csv(bleached.acoef,"LP.bleached.acoef.csv")
write.csv(filterprops.nox,"LP.filterprops.nox.csv")

```

## 2	PREPROCESSING AT THE REPLICATE LEVEL

* The different fractions are calculated ( _a_^E^Phy extracted pigments, 
_a_^E^NAP unextracted material, and _a_^U^ unextracted bleached material).  
* A Savitzky-Golay filter was applied to the fractions to allow the 
continuum removal function to be effective in spectral regions with low 
signal/noise ratio.  
* A correction for the unextracted chl a in  _a_^U^ was performed as:
 + 1) Red peak isolation 650-750nm in  _a_^U^, with height at peak λ.
+ 2) Load metoh pure chlorophyll a spectrum from Taniguchi and Lindsey 2017  
database (<https://doi.org/10.1111/php.13319>)
+ 3) Shift pure chla spectra and multiply peak height for peak _a_^U^.
* Subtract unextracted chlorophyll a  from bleached pigments spectra.  
* Continuum removal only on phycobilins absorbing spectral range.  

First, load the matrices and filter properties
```{r, eval=FALSE}
#---Step 2 ---- Preprocessing at the replicate level-----
#-------Step 2.1) ---- Raw data loading and preparation----
if (!exists("wholepabs.acoef") ) {wholepabs.acoef.uv=
  read.csv("LP.wholepabs.acoef.csv")[,-1]
} else {wholepabs.acoef.uv=wholepabs.acoef}
  if( !exists("extracted.acoef") )
  {extracted.acoef.uv=read.csv("LP.extracted.acoef.csv")[,-1]
} else {extracted.acoef.uv=extracted.acoef}
if( !exists("bleached.acoef") )
{bleached.acoef.uv=read.csv("LP.bleached.acoef.csv")[,-1]
} else {bleached.acoef.uv=bleached.acoef}
if( !exists("filterprops.nox") )
{filterprops.nox=read.csv("LP.filterprops.nox.csv")[,-1]}

library(dplyr);library(data.table);library(reshape);library(hydroGOF)
library(inspectr);library(IDPmisc)
#visible fraction
wholepabs.acoef=wholepabs.acoef.uv[1:401,]
extracted.acoef=extracted.acoef.uv[1:401,]
bleached.acoef=bleached.acoef.uv[1:401,]
#-------Step 2.2) ---- Calculate Differential Fractions ----
extracpigments=(wholepabs.acoef-extracted.acoef)
extracpigments[,1]=extracted.acoef[,1]
bleachpigments=(extracted.acoef-bleached.acoef)
bleachpigments[,1]=extracted.acoef[,1]
totalpigments=(wholepabs.acoef-bleached.acoef)
totalpigments[,1]=extracted.acoef[,1]
```

Savitzky-Golay filter was applied to the fractions to allow the 
continuum removal function to be effective in spectral regions with low 
signal/noise ratio. 
```{r, eval=FALSE}

#--Step 2.3) - Savitzky-Golay filter of extracted and bleached fractions ----
library(signal)
smoothedextracted<-apply(extracted.acoef[,-1],2, FUN=sgolayfilt, 
                         p = 3, n = 31, m = 0, ts = 1)
smoothedextracted=data.frame(extracted.acoef[,1],smoothedextracted)
smoothedextractedpig<-apply(extracpigments[,-1],2, FUN=sgolayfilt, 
                            p = 3, n = 31, m = 0, ts = 1)
smoothedextractedpig=data.frame(extracpigments[,1],smoothedextractedpig)
smoothedbleached<-apply(bleached.acoef[,-1],2, FUN=sgolayfilt, 
                        p = 3, n = 31, m = 0, ts = 1)
smoothedbleached=data.frame(bleached.acoef[,1],smoothedbleached)
smoothedbleachpig<-apply(bleachpigments[,-1],2, FUN=sgolayfilt, 
                         p = 3, n = 31, m = 0, ts = 1)
smoothedbleachpig=data.frame(bleachpigments[,1],smoothedbleachpig)
```
Correction for the unextracted chl a in  _a_^U^ was performed as:  

 + 1) Red peak isolation 650-750nm in  _a_^U^, with height at peak λ.
```{r, eval=FALSE}
#--Step 2.4)  - Correcting the unextracted chl a in bleached pigments ----
#--Step 2.4.a red peak 650-750nm in bleached pigments, height at peak lambda----
bleachedcremov650=apply(smoothedbleachpig[-(152:401),-1],
 2,FUN=continuum_removal,wl=as.numeric(bleachpigments[-(152:401),1]), upper=F)
bleachedcremoval650=data.frame(cbind(smoothedbleachpig[-(152:401),1],
                                     -bleachedcremov650))
bleachedcremoval650[,1]=extracted.acoef[-(152:401),1]
names(bleachedcremoval650)[1]=names(extracted.acoef)[1]
ceros=data.frame(matrix(0L, nrow= (401-151), ncol = ncol(bleachedcremoval650)))  
names(ceros)=names(bleachedcremoval650)
ceros[,1]=extracted.acoef[152:401,1]
cr.bleach.pigments650=rbind(bleachedcremoval650,ceros)
cr.bleach.pigments650.row=apply(cr.bleach.pigments650, 2, which.max) #peak row 
cr.bleach.pigments650.lambda=apply(cr.bleach.pigments650, 2, which.max) # peak 
cr.bleach.pigments650.max=apply(cr.bleach.pigments650, 2, max) #peak height
cr.bleach.pigments650.peak=data.frame(cr.bleach.pigments650.row,
  cr.bleach.pigments650.lambda,cr.bleach.pigments650.max)
#change peak lambda per wavelength
cr.bleach.pigments650.peak[,2]=extracted.acoef[cr.bleach.pigments650.lambda,1]  
#if lambda >700, make value 0
cr.bleach.pigments650.peak[,3] = ifelse(cr.bleach.pigments650.peak[,2]>700,0,cr.bleach.pigments650.peak[,3]) 
#if lambda >700, make value
cr.bleach.pigments650.peak[,1] = ifelse(cr.bleach.pigments650.peak[,2]>700,NA,cr.bleach.pigments650.peak[,1]) 
cr.bleach.pigments650.peak.pr=cr.bleach.pigments650.peak
#NA bleached pigments
cr.bleach.pigments650.peak.pr[1,]=names(cr.bleach.pigments650.peak.pr)
cr.bleach.pigments650.peak.pr=data.frame(t(cr.bleach.pigments650.peak.pr))
names(cr.bleach.pigments650.peak.pr)=names(filterprops.nox)
filterprops.nox=rbind(filterprops.nox,cr.bleach.pigments650.peak.pr)
```
+ 2) Load metoh pure chlorophyll a spectrum from Taniguchi and Lindsey 2017  
database (<https://doi.org/10.1111/php.13319>)
```{r, eval=FALSE}
#---------Step 2.4.b Load metoh pure chlorophyll a spectrum ----  
#PhotochemCAD 3 Taniguchi et al., 2017 database
pure.chla=read.delim("CHL008_Chl a, MeOH (Kobayashi, 2013).abs.txt")
pure.chla.re=prospectr::resample(
  X=pure.chla$CHL008_Chl.a..MeOH..Kobayashi..2013..abs, 
 wav=pure.chla$Wavelength..nm., new.wav=extracted.acoef$Wave.length, 
 interpol = "linear")
pure.chla.re[is.na(pure.chla.re)]<-0
pure.chla=data.frame(nm=extracted.acoef$Wave.length,abs=pure.chla.re)
pure.chla.df=extracted.acoef
pure.chla.df[,-1]<-pure.chla$abs
#max at 135, 666 nm

pure.chla.cr=apply(pure.chla.df[pure.chla.df[,1]>650&pure.chla.df[,1]<750 ,-1],
 2,FUN=continuum_removal,wl=
as.numeric(pure.chla.df[pure.chla.df[,1]>650&pure.chla.df[,1]<750 ,1]), upper=F)
pure.chla.cr=data.frame(
cbind(pure.chla.df[pure.chla.df[,1]>650&pure.chla.df[,1]<750 ,1],-pure.chla.cr))
zeros=data.frame(
  matrix(0L, nrow= nrow(pure.chla.df), ncol = ncol(pure.chla.df)))
names(zeros)=names(pure.chla.df)
zeros[,1]=pure.chla.df[,1]
zeros[match(pure.chla.cr[,1],zeros[,1]),]<-pure.chla.cr
pure.chla.cr<-zeros
pure.chla.cr.factor=data.frame(apply(pure.chla.cr,2,FUN = max)/
                                 apply(pure.chla.df,2,FUN = max))[2,1]
```
+ 3) Shift pure chla spectra and multiply peak height for peak _a_^U^.
```{r, eval=FALSE}
#--Step 2.4.c--Shift chla spectra and multiply peak for bleached peak height----
#multiply the extracted pigments for the ratio of red peaks extracted/bleached
#shift the extracted pigments for the difference in peaks extracted-bleached
lambdashift <- function(x, n){ if (is.na(n)) {rep(NA, length(x))} 
  else if (n>0){ c(rep(0, n),x[(seq(length(x)-n))]) }
  else if (n<0) {c(x[-(seq(-n))],rep(0, -n)) }
  else  {(x) }}
un.extracpigments.new=data.frame((mapply(function(x,n) x*n, x=(pure.chla.df/pure.chla.cr.factor),n=(cr.bleach.pigments650.peak[,3]))))
shifted.un.extracpigments.new=data.frame(mapply(FUN=lambdashift,
                                                (un.extracpigments.new),n=(cr.bleach.pigments650.peak[,1]-135)))
shifted.un.extracpigments.new[,1]=extracpigments[,1]
shifted.un.extracpigments.new=apply(shifted.un.extracpigments.new,2,
 FUN= function (x) if (sum(!is.na(x))==0) {rep(0,length(x))} else {x})
```
Subtract unextracted chlorophyll a  from bleached pigments spectra.  
```{r, eval=FALSE}
#--Step 2.5 -Subtract unextracted chlorophyll from bleached pigments spectra----
true.bleached.new=(smoothedbleachpig-shifted.un.extracpigments.new)
true.bleached.new[,1]=extracpigments[,1]
```
Continuum removal only on phycobilins absorbing spectral range.  
```{r, eval=FALSE}
#--Step 2.6 - Null point correction and continuum removal  on phycobilins ----
#smoothing filter
library(signal)
#continuum removal
cr.true.bleached.new=apply(true.bleached.new[51:351,-1],2,
 FUN=continuum_removal,wl=as.numeric(true.bleached.new[51:351,1]), upper=F)
napfree.bleached.new=data.frame(cbind(true.bleached.new[51:351,1],
                                      -cr.true.bleached.new))
ceros.a=data.frame(matrix(0L, nrow= (51-1), ncol = ncol(true.bleached.new)))  
ceros.b=data.frame(matrix(0L, nrow= (401-351), ncol = ncol(true.bleached.new)))  
names(ceros.a)=c("V1",names(true.bleached.new)[-1])
names(ceros.b)=c("V1",names(true.bleached.new)[-1])
ceros.a[,1]=true.bleached.new[1:50,1];ceros.b[,1]=true.bleached.new[352:401,1]
napfree.bleached.new=rbind(ceros.a,napfree.bleached.new,ceros.b)#yesssssss
# changed smoothed.true.bleached[,-1] for true.bleached[,-1]
bleached.nap.new=data.frame(cbind(true.bleached.new[,1],
                                  true.bleached.new[,-1]+cr.true.bleached.new))
#2021.04.18 reverse continuum removal only negatively curved  spectra
true.bleached.new.rev=t(t(as.matrix(true.bleached.new))-
                          apply(true.bleached.new[1:50,],2,FUN=mean))
true.bleached.new.rev[true.bleached.new.rev<0] <-0
napfree.bleached.new.rev=rbind(ceros.a[,],
cbind(true.bleached.new[51:351,1],-apply(true.bleached.new.rev[51:351,-1],2,
 FUN=continuum_removal,wl=as.numeric(true.bleached.new[51:351,1]), upper=F)),
 ceros.b[,])
phycobilins=napfree.bleached.new.rev #V2, better
library(tidyr);library(dplyr);library(magrittr);library(tibble)
```
Recalculate final fractions
```{r, eval=FALSE}
#----- Step 2.8 ---- Final fractions calculations ----
#---------Step 2.8.a---- aNAP and aPhy calculations ----
NAP=extracted.acoef[,]-shifted.un.extracpigments.new[,]-phycobilins[,]
PIGMENTS=wholepabs.acoef[,]-NAP
#---------Step 2.8.b---- Whole PABS 750-800 nm transfer to NAP ----
NIR=as.data.frame(t(apply(PIGMENTS[1:50,],2,FUN=mean)))
NIR[1:nrow(PIGMENTS),]=NIR[1,]
FINAL.PIGMENTS=PIGMENTS[,]-NIR;FINAL.PIGMENTS[,1]=extracted.acoef[,1]
FINAL.NAP=NAP[,]+NIR;FINAL.NAP[,1]=extracted.acoef[,1]
PHYCOBILINS=phycobilins[,]
#------Step 2.9 ---- Export ----
write.csv(FINAL.NAP,"FINAL.NAP.csv")
write.csv(FINAL.PIGMENTS,"FINAL.PIGMENTS.csv")
write.csv(PHYCOBILINS,"PHYCOBILINS.csv")

```

## 3	PREPROCESSING AT THE LAKE LEVEL

* The lake means and coefficient of variation of the different fractions 
are calculated for each lake ( _a_^Corr^Phy corrected pigments including 
phycobilins, _a_^Corr^NAP corrected NAP, and _a_^PB^ only the estimated 
phycobilins) 

First, load the matrices and filter properties
``` {r, eval=FALSE}
#---Step 3 ---- Preprocessing per lake-----
library(tidyr);library(dplyr);library(magrittr);library(tibble)
if (!exists("wholepabs.acoef") ) {wholepabs.acoef=
  read.csv("LP.wholepabs.acoef.csv")[,-1]
} else {wholepabs.acoef.uv=wholepabs.acoef}
if (!exists("FINAL.NAP") ) {FINAL.NAP=
  read.csv("FINAL.NAP.csv")[,-1]} 
  if( !exists("FINAL.PIGMENTS") )
  {FINAL.PIGMENTS=read.csv("FINAL.PIGMENTS.csv")[,-1]} 
if( !exists("PHYCOBILINS") )
{PHYCOBILINS=read.csv("PHYCOBILINS.csv")[,-1]}
if( !exists("filterprops.nox") )
  {filterprops.nox=read.csv("LP.filterprops.nox.csv")[,-1]}

true.wholepabs=wholepabs.acoef
true.NAP=FINAL.NAP
true.pigments=FINAL.PIGMENTS
true.phycobilins=PHYCOBILINS

names(true.phycobilins)[1]=names(wholepabs.acoef)[1]
rownames(filterprops.nox)=filterprops.nox[,1]
filterprops.nox.perlake=filterprops.nox%>%t%>%as.data.frame%>%
  group_by(lake_id)%>%summarise_all(funs(paste),collapse="|")%>%t%>%
  as.data.frame
filterprops.nox.perlake=filterprops.nox.perlake[,
  filterprops.nox.perlake[1,]!="lake_id"]
``` 
Lake means for total particulate absorption.
``` {r, eval=FALSE}
#------Step 3.1 ---- PABS Average per lake-----
true.wholepabs=rbind(as.matrix(filterprops.nox)[2,],true.wholepabs)
rownames(true.wholepabs)<-NULL
true.wholepabs.perlake=true.wholepabs%>%column_to_rownames("Wave.length")%>%
  t%>%as.data.frame%>%mutate_at(vars(-lake_id), as.numeric)%>%group_by(lake_id)%>%summarise_all(funs(mean))%>%t%>%as.data.frame
names(true.wholepabs.perlake)=true.wholepabs.perlake[1,]
true.wholepabs.perlake=true.wholepabs.perlake[-1,]
true.wholepabs.perlake=cbind(true.wholepabs[-1,1],true.wholepabs.perlake)
true.wholepabs.perlake=apply(true.wholepabs.perlake,2,FUN=as.numeric)
``` 
Lake means for _a_^Corr^Phy corrected pigments.
``` {r, eval=FALSE}
#------Step 3.2 ---- NAP Average per lake-----
true.NAP=rbind(as.matrix(filterprops.nox)[2,],true.NAP)
rownames(true.NAP)<-NULL
true.NAP.perlake=true.NAP%>%column_to_rownames("Wave.length")%>%t%>%
  as.data.frame%>%mutate_at(vars(-lake_id), as.numeric)%>%group_by(lake_id)%>%summarise_all(funs(mean))%>%t%>%as.data.frame
names(true.NAP.perlake)=true.NAP.perlake[1,]
true.NAP.perlake=true.NAP.perlake[-1,]
true.NAP.perlake=cbind(true.NAP[-1,1],true.NAP.perlake)
true.NAP.perlake=apply(true.NAP.perlake,2,FUN=as.numeric)
``` 
Lake means for _a_^Corr^NAP corrected non algal material.
``` {r, eval=FALSE}
#------Step 3.3 ---- Pigments Average per lake-----
true.pigments=rbind(as.matrix(filterprops.nox)[2,],true.pigments)
rownames(true.pigments)<-NULL
true.pigments.perlake=true.pigments%>%column_to_rownames("Wave.length")%>%
  t%>%as.data.frame%>%mutate_at(vars(-lake_id), as.numeric)%>%group_by(lake_id)%>%summarise_all(funs(mean))%>%t%>%as.data.frame
names(true.pigments.perlake)=true.pigments.perlake[1,]
true.pigments.perlake=true.pigments.perlake[-1,]
true.pigments.perlake=cbind(true.pigments[-1,1],true.pigments.perlake)
true.pigments.perlake=apply(true.pigments.perlake,2,FUN=as.numeric)
```
Lake means for  _a_^PB^ (only the estimated phycobilins) 
``` {r, eval=FALSE}
#------Step 3.4 ---- Phycobilins Average per lake-----
true.phycobilins=rbind(as.matrix(filterprops.nox)[2,],true.phycobilins)
rownames(true.phycobilins)<-NULL
true.phycobilins.perlake=true.phycobilins%>%column_to_rownames("Wave.length")%>%
  t%>%as.data.frame%>%mutate_at(vars(-lake_id), as.numeric)%>%group_by(lake_id)%>%summarise_all(funs(mean))%>%t%>%as.data.frame
names(true.phycobilins.perlake)=true.phycobilins.perlake[1,]
true.phycobilins.perlake=true.phycobilins.perlake[-1,]
true.phycobilins.perlake=cbind(true.phycobilins[-1,1],true.phycobilins.perlake)
true.phycobilins.perlake=apply(true.phycobilins.perlake,2,FUN=as.numeric)
```
The coefficient of variation is calculated for each fraction
``` {r, eval=FALSE}
#------Step 3.5 ---- Standard deviations per lake-----
true.wholepabs.perlake.sd=true.wholepabs%>%column_to_rownames("Wave.length")%>%
  t%>%as.data.frame%>%mutate_at(vars(-lake_id), as.numeric)%>%group_by(lake_id)%>%summarise_all(funs(sd))%>%t%>%as.data.frame
names(true.wholepabs.perlake.sd)=true.wholepabs.perlake.sd[1,]
true.wholepabs.perlake.sd=true.wholepabs.perlake.sd[-1,]
true.wholepabs.perlake.sd=cbind(true.wholepabs[-1,1],true.wholepabs.perlake.sd)
true.wholepabs.perlake.sd=apply(true.wholepabs.perlake.sd,2,FUN=as.numeric)
#
true.NAP.perlake.sd=true.NAP%>%column_to_rownames("Wave.length")%>%t%>%
  as.data.frame%>%mutate_at(vars(-lake_id), as.numeric)%>%group_by(lake_id)%>%summarise_all(funs(sd))%>%t%>%as.data.frame
names(true.NAP.perlake.sd)=true.NAP.perlake.sd[1,]
true.NAP.perlake.sd=true.NAP.perlake.sd[-1,]
true.NAP.perlake.sd=cbind(true.NAP[-1,1],true.NAP.perlake.sd)
true.NAP.perlake.sd=apply(true.NAP.perlake.sd,2,FUN=as.numeric)
#
true.pigments.perlake.sd=true.pigments%>%column_to_rownames("Wave.length")%>%
  t%>%as.data.frame%>%mutate_at(vars(-lake_id), as.numeric)%>%group_by(lake_id)%>%summarise_all(funs(sd))%>%t%>%as.data.frame
names(true.pigments.perlake.sd)=true.pigments.perlake.sd[1,]
true.pigments.perlake.sd=true.pigments.perlake.sd[-1,]
true.pigments.perlake.sd=cbind(true.pigments[-1,1],true.pigments.perlake.sd)
true.pigments.perlake.sd=apply(true.pigments.perlake.sd,2,FUN=as.numeric)
#
true.phycobilins.perlake.sd=true.phycobilins%>%
  column_to_rownames("Wave.length")%>%
  t%>%as.data.frame%>%mutate_at(vars(-lake_id), as.numeric)%>%group_by(lake_id)%>%summarise_all(funs(sd))%>%t%>%as.data.frame
names(true.phycobilins.perlake.sd)=true.phycobilins.perlake.sd[1,]
true.phycobilins.perlake.sd=true.phycobilins.perlake.sd[-1,]
true.phycobilins.perlake.sd=cbind(true.phycobilins[-1,1],
                                  true.phycobilins.perlake.sd)
true.phycobilins.perlake.sd=apply(true.phycobilins.perlake.sd,2,FUN=as.numeric)
#------Step 3.6 ---- Coef of variation per lake-----
true.wholepabs.perlake.cvar=100*true.wholepabs.perlake.sd/true.wholepabs.perlake
true.NAP.perlake.cvar=100*true.NAP.perlake.sd/true.NAP.perlake
true.pigments.perlake.cvar=100*true.pigments.perlake.sd/
  true.pigments.perlake
true.phycobilins.perlake.cvar=100*true.phycobilins.perlake.sd/
  true.phycobilins.perlake

```
Export the calculated datasets
``` {r, eval=FALSE}
#------Step 3.7 ---- Export-----
write.csv(true.wholepabs.perlake,"true.wholepabs.perlake.csv")
write.csv(true.NAP.perlake,"true.NAP.perlake.csv")
write.csv(true.pigments.perlake,"true.pigments.perlake.csv")
write.csv(true.phycobilins.perlake,"true.phycobilins.perlake.csv")

write.csv(true.wholepabs.perlake.cvar,"true.wholepabs.perlake.cvar.csv")
write.csv(true.NAP.perlake.cvar,"true.NAP.perlake.cvar.csv")
write.csv(true.pigments.perlake.cvar,"true.pigments.perlake.cvar.csv")
write.csv(true.phycobilins.perlake.cvar,"true.phycobilins.perlake.cvar.csv")


```